### 2020-04-07 0_abstract.tex
Apply Alex's comments.

\begin{abstract}
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench lowers the barrier to benchmarking multiple programming systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: every benchmark constructed with Task Bench
runs on every Task Bench implementation. Furthermore,
Task Bench's parameterization enables a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

>>>>

\begin{abstract}
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench dramatically lowers the barrier to benchmarking and comparing multiple programming systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: every benchmark constructed with Task Bench
runs on every Task Bench implementation. Furthermore,
Task Bench's parameterization enables a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

### 2020-04-07 1_introduction.tex
Apply Alex's comments.

We conduct a comprehensive study of all 15 Task Bench implementations on
up to 256 Haswell nodes of the Cori supercomputer~\cite{Cori}.
%% :
%% Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP, CUDA),
%% OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
%% PaRSEC \cite{PARSEC13}, Realm \cite{Realm14}, Regent \cite{Regent15},
%% Spark \cite{Spark10}, StarPU \cite{StarPU11},
%% Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
%% X10 \cite{X1005}.
Using METG, we find that a number of factors---node
count, accelerators, and complex dependencies, among
others---individually or in combination contribute to an order of
magnitude or greater increase in METG, even in the most efficient
systems. While some systems can achieve sub-microsecond METG(50\%) in
best-case scenarios, we show that a more realistic
bound for running nearly any application at scale is 100~\textmu{}s with
current technologies. Our study includes several asynchronous systems
which are designed to provide benefits such as overlapped computation
and communication. While small-scale benchmarks of these systems
suffer from increased overhead, we find that the benefits of these systems become
tangible at scale (provided the runtime overhead doesn't increase
beyond about 100~\textmu{}s).

@@

Beyond comparative study, the ability to explore a large configuration
space also enables the discovery of bugs in the underlying systems. We
found five performance issues, ranging from communication efficiency
(Chapel, Realm), to the efficiency of task pruning, analysis and
constant folding (PaRSEC, Dask and TensorFlow). Three have been fixed
and all have been acknowledged by the developers of the respective
systems. (All were either fixed or worked around in our experiments.)
In some cases these correspond to order of magnitude (or even
asymptotic) improvements in the performance of the underlying
systems---benefits which apply well beyond Task Bench to all classes
of applications.

The paper is organized as follows: Section~\ref{sec:task-bench}
describes the Task Bench design. Section~\ref{sec:implementation}
discusses implementations in 15 systems.  Section~\ref{sec:metg}
defines METG and its relationship to quantities of interest to
application developers.  Section~\ref{sec:evaluation} provides a
comprehensive evaluation on Cori. Section~\ref{sec:related-work} relates to
previous efforts; Section~\ref{sec:conclusion} concludes.

>>>>

We conduct a comprehensive study of all 15 Task Bench implementations on
up to 256 Haswell nodes of the Cori supercomputer~\cite{Cori}.
%% :
%% Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP, CUDA),
%% OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
%% PaRSEC \cite{PARSEC13}, Realm \cite{Realm14}, Regent \cite{Regent15},
%% Spark \cite{Spark10}, StarPU \cite{StarPU11},
%% Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
%% X10 \cite{X1005}.
Using METG, we find that a number of factors---node
count, accelerators, and complex dependencies, among
others---individually or in combination contribute to an order of
magnitude or greater increase in METG, even in the most efficient
systems. While some systems can achieve sub-microsecond METG(50\%) in
best-case scenarios, we show that a more realistic
bound for running nearly any application at scale is 100~\textmu{}s with
current technologies. Our study includes several asynchronous systems
designed to provide benefits such as overlapped computation
and communication. While small-scale benchmarks of these systems
suffer from increased overhead, we find that the benefits of these systems become
tangible at scale (provided the runtime overhead doesn't increase
beyond about 100~\textmu{}s per task).

@@

Beyond comparative study, the ability to explore a large configuration
space also enables the discovery of bugs in the underlying systems. We
found five performance issues, ranging from communication efficiency
(Chapel, Realm), to the efficiency of task pruning, analysis and
constant folding (PaRSEC, Dask and TensorFlow). Three have been fixed
and all have been acknowledged by the developers of the respective
systems. (All were either fixed or worked around in our experiments.)
In some cases these correspond to order of magnitude (or even
asymptotic) improvements in the performance of the underlying
systems---benefits which apply well beyond Task Bench to all classes
of applications. The bugs are described in more detail in Section~\ref{sec:case-study}.

The paper is organized as follows: Section~\ref{sec:task-bench}
describes the Task Bench design. Section~\ref{sec:implementation}
discusses implementations in 15 systems.  Section~\ref{sec:metg}
defines METG and its relationship to quantities of interest to
application developers.  Section~\ref{sec:evaluation} provides a
comprehensive evaluation on Cori. Section~\ref{sec:case-study} describes bugs found with Task Bench. Section~\ref{sec:related-work} relates to
previous efforts; Section~\ref{sec:conclusion} concludes.

### 2020-04-07 4_metg.tex
Apply Alex's comments.

The choice of 50\% is also a parameter and is not fundamental to
METG. We use 50\% in our studies to avoid pathologies
associated with lower efficiency thresholds (see
Section~\ref{subsec:peak-performance-and-efficiency}), and also
because it aligns with what we observe in practice. For example, in one supercomputer center's application for new projects, users are instructed to "select the
most parallel efficient job size (ratio of benchmark speed-up
vs. linear speed-up above 50\%)" \cite{CSCSReport}. This corresponds
to METG(50\%).

@@

Figures~\ref{fig:flops-mpi} and \ref{fig:efficiency-mpi} show how METG is
calculated. The application, in this case an MPI implementation of the
Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
run on a single Haswell node of Cori with a problem size large enough
that runtime is dominated by the application's inner loops. This confirms
that the application is properly configured and that the
efficiency metric is achievable. The problem
size is then repeatedly reduced while maintaining exactly the same hardware and software
configuration (in particular, the same number of nodes and tasks). The
expectation is that as problem size shrinks,
performance will begin to drop and eventually approach zero, as shown in Figure~\ref{fig:flops-mpi}. Systems
with lower runtime overheads maintain higher performance at smaller
problem sizes compared to systems with higher overheads.

@@

METG(50\%) is the intersection of this curve with 50\% efficiency, as
shown by the red, dashed lines in Figures~\ref{fig:flops-mpi} and
\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
task granularity of
4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s in this
configuration. We use 50\% because that is generally an acceptable level
of efficiency in practice, and values above 90\% can misrepresent the
performance of some systems (see
Section~\ref{subsec:peak-performance-and-efficiency}).

>>>>

The choice of 50\% is also a parameter and is not fundamental to
METG. We use 50\% in our studies to avoid pathologies
associated with lower efficiency thresholds (see
Section~\ref{subsec:peak-performance-and-efficiency}), and also
because it aligns with what we observe in practice. For example, in one supercomputer center's application for new projects, users are instructed to ``select the
most parallel efficient job size (ratio of benchmark speed-up
vs. linear speed-up above 50\%)'' \cite{CSCSReport}, which corresponds
to METG(50\%).

@@

Figures~\ref{fig:flops-mpi} and \ref{fig:efficiency-mpi} show how METG is
calculated. The application, in this case an MPI implementation of the
Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
run on a single Haswell node of Cori with a problem size large enough
that runtime is dominated by the application's inner loops. This result confirms
that the application is properly configured and that the
efficiency metric is achievable. The problem
size is then repeatedly reduced while maintaining exactly the same hardware and software
configuration (in particular, the same number of nodes and tasks). The
expectation is that as problem size shrinks,
performance will begin to drop and eventually approach zero, as shown in Figure~\ref{fig:flops-mpi}. Systems
with lower runtime overheads maintain higher performance at smaller
problem sizes compared to systems with higher overheads.

@@

METG(50\%) is the intersection of this curve with 50\% efficiency, as
shown by the red, dashed lines in Figures~\ref{fig:flops-mpi} and
\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
task granularity of
4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s in this
configuration. %% We use 50\% because that is generally an acceptable level
%% of efficiency in practice, and values above 90\% can misrepresent the
%% performance of some systems (see
%% Section~\ref{subsec:peak-performance-and-efficiency}).

### 2020-04-07 5_evaluation.tex
Fiddle with figure placement, caption size, prune communication experiment.
Add note on omitted load imbalance results.
Apply Alex's comments.

@@

Figure~\ref{fig:flops} shows the FLOP/s achieved with a compute-bound
kernel with varying problem sizes (simulated by running the kernel for varying numbers of iterations). This is the full version of
Figure~\ref{fig:flops-mpi}. Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOP/s of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency, we use our empirically determined number as
the baseline for 100\% efficiency.

@@

One question when considering different programming
systems is: how much overhead does the system add? This question is tricky to answer directly because some systems introduce
overhead \emph{inline} (i.e., by running system internal processes on
the same cores as application tasks), while other systems introduce
overhead \emph{out-of-line} (i.e., by dedicating one or more cores
solely to runtime use). Some systems, like Charm++, Realm, and Regent,
support both configurations.

@@

\subsection{Load Imbalance}

@@

The remaining differences are due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities. We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave analysis of persistent load imbalance to future work.

>>>>

\emph{Due to space constraints we omit our load balancing results, which will be included in the revision and final paper.}

@@

Figure~\ref{fig:flops}, which is the full version of
Figure~\ref{fig:flops-mpi}, shows the FLOP/s achieved with a compute-bound
kernel with varying problem sizes (simulated by running the kernel for varying numbers of iterations). Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOP/s of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency, we use our empirically determined number as
the baseline for 100\% efficiency.

@@

One question when considering different programming
systems is: How much overhead does the system add? This question is tricky to answer directly because some systems introduce
overhead \emph{inline} (i.e., by running system internal processes on
the same cores as application tasks), while other systems introduce
overhead \emph{out-of-line} (i.e., by dedicating one or more cores
solely to runtime use). Some systems, like Charm++, Realm, and Regent,
support both configurations.

@@

\zap{
\subsection{Load Imbalance}

@@

The remaining differences are due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities. We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave analysis of persistent load imbalance to future work.
} % zap

### 2020-04-07 8_case_study.tex
Apply Alex's comments.

Task Bench is useful not only as a tool for evaluating programming
system performance, but also for discovering potential areas of
improvements. During the development of Task Bench, we identified a
number of performance issues in the underlying programming
systems. These discoveries were possible because of the flexibility of
Task Bench, and our ability to rapidly run new experiments with a
variety of application scenarios. All issues were reported to and
acknowledged by the respective system's developers, and are either
fixed or worked around in our experiments. For brevity, we describe
one salient example below and summarize the others briefly.

Realm and Chapel both use DMA subsystems that are optimized for large
copies. Early Task Bench experiments revealed high METG values that
were diagnosed by the Realm/Chapel developers as overhead due to the
cost of scheduling small copies. Subsequent improvements in Realm and
Chapel improved small copy overheads (and thus METG) by over an order
of magnitude in the case of Realm, and by $2\times$ in the case of
Chapel. These improvements affect any application where fine-grained
tasks are used, which is particularly relevant in strong scaling
regimes when running on large numbers of nodes.

@@

Chapel uses a naive round-robin scheduler by default, which can lead
to unexpected load balance issues because certain language features
(such as remote array assignment) implicitly generate tasks. This
resulted in poor peak performance, even at large task granularities,
which in some cases made it impossible to measure METG (because peak
efficiency did not exceed 50\%). Chapel's other schedulers add
overhead, resulting in higher METGs. We worked around this with
\lstinline{serial}.

>>>>

Task Bench is useful not only as a tool for evaluating programming
system performance, but also for discovering potential areas for
improvement. During the development of Task Bench, we identified a
number of performance issues in the underlying programming
systems. These discoveries were possible because of the flexibility of
Task Bench, and our ability to rapidly run new experiments with a
variety of application scenarios. All issues were reported to and
acknowledged by the respective system's developers, and are either
fixed or worked around in our experiments. For brevity, we describe
one salient example below and summarize the others briefly.

Realm and Chapel both use DMA subsystems that are optimized for large
copies. Early Task Bench experiments revealed high METG values that
were diagnosed by the Realm/Chapel developers as overhead due to the
cost of scheduling small copies. Subsequent improvements in Realm and
Chapel improved small copy overheads (and thus METG) by over an order
of magnitude in the case of Realm, and by $2\times$ in the case of
Chapel. These improvements affect any application where fine-grained
data movement is needed, which is particularly relevant in strong scaling
regimes when running on large numbers of nodes.

@@

Chapel uses a naive round-robin scheduler by default, which can lead
to unexpected load balance issues because certain language features
(such as remote array assignment) implicitly generate tasks. This
resulted in poor peak performance, even at large task granularities,
which in some cases made it impossible to measure METG (because peak
efficiency did not exceed 50\%). Chapel's other schedulers add
overhead, resulting in higher METGs. We worked around this issue with
\lstinline{serial}.
